---
title             : "Non-word repetition in Yélî Dnye"
shorttitle        : "NWR in Yélî Dnye"

author: 
  - name          : "Alejandrina Cristia"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "29, rue d’Ulm, 75005 Paris, France"
    email         : "alecristia@gmail.com"
  - name          : "Marisa Casillas"
    affiliation   : "2"
    address       : "Wundtlaan 1, 6525 XD, Nijmegen, The Netherlands"
    email         : "marisa.casillas@mpi.nl"

affiliation:
  - id            : "1"
    institution   : "Laboratoire de Sciences Cognitives et de Psycholinguistique, Département d'Etudes cognitives, ENS, EHESS, CNRS, PSL University"
  - id            : "2"
    institution   : "Max Planck Institute for Psycholinguistics"

authornote: |
  All data are made available in a repository in the Open Science Framework. AC acknowledges the support of the Agence Nationale de la Recherche (ANR-17-CE28-0007 LangAge, ANR-16-DATA-0004, ANR-14-CE30-0003, ANR-17-EURE-0017); and the J. S. McDonnell Foundation Understanding Human Cognition Scholar Award. 
   

  
abstract: >
   In nonword repetition (NWR) studies, participants are presented auditorily with an item that is phonologically legal but lexically meaningless in the local language. Accuracy is thought to reflect long-term phonological knowledge as well as online phonological working memory and flexible production patterns.In this study, we report on NWR results among children learning Yêly Dnyé, an isolate spoken in Rossel Island, PNG, with an unusually dense phonological inventory. This study contributes to three lines of research. First, we document that non-word items containing typologically rare sounds are repeated accurately less often that non-words containing more common sounds. Second, we document rather weak effects of item length, contributing to other research suggesting that length effects may be language-specific. Third, we do not find strong individual variation effects in this population, contrary to previous results documenting strong age-related effects. Together, these data provide a unique view of online phonological processing in a seldom-studied language, and contribute to both typological and language acquisition research.




keywords          : "phonology, non-word repetition, development"
wordcount         : "~xxx words"

bibliography      : ["nwrbib.bib"]

figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes

lang              : "english"
class             : "man"
output            : papaja::apa6_pdf
---



```{r setup, include=FALSE}
REDOALL=FALSE #use true if you want to recalculate the final-data and other data sheets

#Sys.setlocale("LC_CTYPE", "en_US_POSIX.UTF-8")
#Sys.setlocale("LANG", "en_US_POSIX.UTF-8")
Sys.setlocale(category = "LC_ALL", locale = "en_SE.utf8")

knitr::opts_chunk$set(warning = FALSE, message = FALSE, cache = FALSE, fig.pos = "T")
require("papaja")
library(kableExtra)
library(gdata)
library(ggplot2)
library(lme4)

correspondances=matrix( #list correspondances always by pairs, orthography then phonology
  c(
    #nasal  to unichar
    ":ii","ì",
    ":ee","è",
    ":aa","à",
    ":êê","3",
    ":ââ","6",
    ":uu","ù",
    ":oo","Ò",
    #long to unichar
    "ii","ï",
    "ee","ë",
    "éé","Ë",
    "áá","Ä",
    "aa","ä",
    "îî","Ï",
    "êê","2",
    "ââ","5",
    "uu","ü",
    "óó","ö",
    "oo","Ö",
    #short nasal
    ":i","1",
    ":e","4",
    ":a","ã",
    ":ê","6",
    ":â","7",
    ":u","8",
    ":o","9",
    #consonants, veeeery dirty!
    "'n","ń",
    "tp",":",
    "dp","/",
    "kp","P",
    "ngm","G",
    "ńm","M",
    "ng","N",
    "mb","b",
    "nd","D",
    "nt","T",
    "nk","K",
    "dn","?",
    "kn","!",
    "vy","V",
    "mt","C",
    "md","J",
    "mg","X",
    "dm","S",
    "km","Z",
    "gh","H",
    "lv","L"
  ),#last item above should not have a comma
  ncol = 2,byrow = T)

```

<!-- REMEMBER!!!! -->
<!-- phonetic characters: -->
<!-- - save excel as utf with save as -->
<!-- - do $ iconv -f utf-16le -t utf-8 stimuli.txt > stimconv.txt -->
<!-- - do $ iconv -f utf-16le -t utf-8 segments.txt > segconv.txt -->


## TODO Middy

- look through this manuscript for bits of text to be written by **mc**
- help with table 1 (formatting & making it appear, because it's broken by the phonological representations)
- do you think Figure 2 should be done with phonetic symbols? If yes, could you try to implement that?
- double check demo -- probably start with the xlsx version that has the children's first names, date of birth, and date of test. I looked again at my local files (still haven't checked the labs' ones) and I'm surprised I have maternal education for more children than I have birth order for... If you correct this, to generate the csv I also round the age, assign age.rounded to reported age (in years) when DOB is missing, remove first names, DOB, DOTest and age.precise before pushing to github
- double check my frequency entries in segments.xlsx using the search function in https://phoible.org/parameters; take a look at http://phoible.github.io/conventions/ in case you see something about double articulation (I didn't find tp or lBj)
- double check phonological implementation of stimuli in stimuli.xlsx 
- for both of the above, if you make any changes, they need to be re-saved as UTF16 txt file, and then converted (see remember above)
- double check also stim_seg_freq -- why are some items still with NA freq?
- consent, "correct", "accuracy", "performance"
- draft discussion


## TODO later
- probably turn summary of types of error into a table; use proportion of correctly produced rather than number; use first attempts rather than the whole data set
- maybe add to the last ana the prevalence (ie out of all the gh which proportion get transformed)
- maybe add analyses of score as a function of frequency (phones, diphones) in MC's corpus? (mc suggested, but I don't think we need it...)


## Introduction

Although infants begin to learn about their native language's phonology within the first year, many studies suggest that in perception and production, in phonetics and phonology, their knowledge continues to develop throughout childhood [e.g., @hazan2000development]. One common task in this line of research is nonword repetition (NWR). In NWR studies, participants are presented auditorily with an item that is phonologically legal but lexically meaningless in the language children are learning. The child should immediately try to say it back without changing anything. Accuracy is thought to reflect long-term phonological knowledge (which allows the child to perceive the item accurately even though it is not a real word they have encountered before) as well as online phonological working memory (to encode the item in the interval between hearing it and saying it back) and flexible production patterns (to produce the item accurately even though it had never been pronounced before). 

NWR has been used to seek answers to a variety of theoretical questions, including what  the links between phonology, working memory, and the lexicon are [@bowey2001nonword], and to assess the extent to which apparent phonological constraints found in the lexicon affect online production [@gallagher2014identity]. It is also frequently used as for applications, notably  as a diagnostic for language delays and disorders [@estes2007differences]. Since non-words can be generated in any language, it has attracted the attention of researchers working in multilingual and linguistically diverse environments,  particularly in Europe [@meir2016disentangling;@is08042009language]. 

<!--M2A: I feel like there's a bit of a jump here from explaining what NWR is and what it's used for to saying we used it in the current study to test three things. We should probably have a paragraph justifying its use in this context, no?-->

In this study, we report on NWR results among children learning Yêly Dnyé, an isolate spoken in Rossel Island, PNG, with an unusually dense phonological inventory.  The present report uses these unique data to contribute to several lines of research. First, we  made sure that some of the non-word items contained typologically rare and/or  challenging sounds, so that we can contribute to the study of whether rare sounds are disadvantaged in perception and/or production, both in terms of overall accuracy of repetition and patterns of mispronunciation. Second, we look at the effects of item length, since previous NWR research has uncovered variable effects. Third, we contribute to the basic NWR literature by contributing an additional data point for children learning a rarely-studied language and culture, and further investigating whether there are structured sources that account for individual variation.

### Intro to the language (**mc**) - please feel free to throw away anything that is not useful!

- complexity in the vowel system
- complexity in the consonant system
- word shapes
- typical word length
- although not the focus of this paper, high use of suppletion in verbal paradigms, other features of language, see Levinson XXX for details

### Intro to the people (**mc**) - please feel free to throw away anything that is not useful!

Little is known about language development in children growing up in Rossel Island, a community of primarily subsistence farmers who tend to reside in close-knitted villages where child care is distributed across many individuals, and who typically speak Yélî Dnyé, a phonologically and lexically complex language. 

- usually monolingual at home
- schooling in English but it starts at age XX, so not relevant here
- however, some use of English due to immigrants & children of immigrants
- children spend a lot of time with other children
- most parents are subsistence farmers
- parental education generally varies between XX and YY

### Brief review of NWR for our purposes

There is some variation in the presentation procedure and structure of items found in previous NWR work. For example, items are often presented orally by the experimenter [@torrington2015non], although an increasing number of studies have turned to playing back the stimuli in order to have greater control of the stability of the presentation [@brandeker2015language]. Additionally, while some studies have used 10-15 non-words, others have employed up to 46 unique items [@piazzalunga2019articulatory]. Often, authors modulate structural complexity, typically measured in terms of item length (measured in number of syllables) and/or syllable structure [open as opposed to closed syllables, @gallon2007non]. 

Previous work seems to avoid difficult sounds, but we felt this was important to represent Yélî Dnye, so we also varied this factor. We designed a relatively large number of items but, aware that this may render the task longer and more tiresome, we split some of the items across children. This allowed us to get information about repetition accuracy of more items. 

Naturally, designing the task in this way may render the study of individual variaiton within the population  more difficult because different children are exposed to different items. However, a review of previous work on individual variation suggested to us that many individual differences effects are relatively small, and would not be detectable with the sample size that we could collect in a given visit. <!--M2self: make sure it's apparent how/why a visit is limited-->

That said, we contribute to the literature by also reporting descriptive analyses of individual variation that could potentially be integrated in meta-analytic efforts. Based on previous work, we looked at potential improvements with age [@vance2005speech;@farmani2018normalization;@kalnak2014nonword], and potential negative effects of bilingual exposure [@brandeker2015language;@meir2016disentangling;@meir2017independent]. Previous work typically finds no significant differences as a function of maternal education [e.g., @farmani2018normalization;@kalnak2014nonword; @meir2017independent] or child gender [@chiat2007preschool]. Although previous research has not often investigated potential effects of birth order on NWR, there is a sizable literature on these effects in other language tasks [@havron2019effect], and therefore we report on those too.



### Research questions

After some preliminary analyses to set the stage, we address the following questions:


- Does the frequency of sounds across languages predict NWR? Are rarer sounds more often substituted by commoner sounds?
- How does score change as a function of item length in number of syllables?
- Is individual variation attributable to child age, sex, birth order, monolingual status, and/or parental education?
<!--M2A: I'm imagining a reviewer who's knowlegeable about phonology and child development but who hasn't encountered NWR before. I think it would be beneficial to provide a concrete example of a nonword and an inexact repetition above so that they can conceive of what these questions imply (analytically) before we get into the methodological weeds.-->


In view of the hypothesis-driven nature of this work, we had considered boosting the interpretational value of this evidence by announcing our analysis plans prior to conducting them. However, we realized that even pre-registering an analysis would be equivocal because we do not have enough power to look at all relationships of interest, and often to detect any of the known effects given their variability across studies. To illustrate this, we portray studies in which children's NWR scores were gathered between 4 and 12 years of age, and reported separately for items that are relatively short (1-2 syllables) versus longer items (3-4 syllables) in Figure 1. Notice that the effect of stimulus length is minuscule among Italian children [@piazzalunga2019articulatory], but considerable among Tsimane' children  [@cristia2020infant], where a drop of 40 percentage points is observed at all ages. A similar difference in NWR scores for short versus long items was observed among Arabic children [@jabere2018xperiment].  Even the effect of age is unstable in this sample. Whereas it is quite clear that children's NWR scores increase in the Italian data, age effects are less stable among Tsimane' children.  Therefore, all analyses here are descriptive and should be considered exploratory.


```{r prevlit-fig, echo = FALSE, ig.pos = "T!", fig.width=8, fig.height=5.5, fig.cap = "NWR scores as a function of age (in years) and item length for comparable studies. Arabic data  from Jaber et al. (2018); Italian data from Piazzalunga et al. (2019); Tsimane' from Cristia et al. (2020); Yélî data from the present study."}

pl<- read.delim("~/Documents/gitrepos/nwryd19/pl-With_yd.tsv")
palette = c( 
             "Arabic 1-2 syll" = "orange",
             "Yélî 3-4 syll"   ="navyblue",
             "Yélî 1-2 syll"  ="lightblue",
             "Italian 3-4 syll"   ="black",
             "Italian 1-2 syll"  ="gray",
             "Arabic 3-4 syll"     = "orange", 
             "Tsimane' 1-2 syll" = "green",
              "Tsimane' 3-4 syll" = "darkgreen")


myplot=ggplot(data = pl, aes(x = Age, y = Score, color = Language))+
  geom_point(aes(size = n, shape = Length))+
  geom_line(aes(linetype=Length)) + 
  theme(text = element_text(size = 16), legend.position="none", axis.text.x=element_text(colour=c("black", "black"))) +
  scale_color_manual(values = palette)+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  scale_x_continuous(name = "Age (years) ", breaks = c(4,8,12))+
  
  scale_y_continuous(name = "NWR Score", breaks = seq(0, 100, 10))+
  scale_size_continuous(range = c(2, 6)) 

myplot

```




## Methods

### Stimuli

Many NWR studies are based on a fixed list of 12-16 items that vary in length between 1 and 4 syllables, often additionally varying syllable complexity and/or cluster presence and complexity, always meeting the condition that they do not mean anything in the target language [e.g., @balladares2016socio; @wilsenach2013phonological]. We kept the same variation in item length and the non-meaningfulness requirement, but we did not vary syllable complexity and clusters because these are vanishingly rare in Yély Dnye. We also increased the number of items an individual child would be tested on, so that a child would get up to 23 items to repeat [note that up to other work has also used 24-30 items: @jabere2018xperiment; @kalnak2014nonword], and we created more items and distributed them across children, so as to increase the coverage, and be able to study more items.

A first list of candidate items was generated in 2018 by selecting simple consonants ("p",	"t",	"d",	"k",	"m",	"n",	"w",	"y") and vowels ("i", "o","u", "a","e") that were combined into consonant-vowel syllables, further sampling the space of 1- to 4-syllable sequences. These candidates were automatically checked against Levinson's 2015 dictionary and removed from consideration if they appeared in the dictionary. The second author presented them orally to three local research assistants, who were asked to repeat them and further say whether they were real words. Any item for which two or more of the assistants reported them having a meaning or some form of association was excluded.

A second list of candidate items was generated in 2019 by selecting complex consonants and systematically crossing them with all the vowels in the Yélî inventory to produce consonant-vowel monosyllables. As before, items were automatically excluded if they appeared in the dictionary. Additionally, since hearing vowel length in monosyllables in isolation is challenging, any item that had a short/long real word neighbor was filtered out. Since the phonology and phonetics of Yélî is still in the process of being described [CITE **mc** please fill in], there could have been undocumented constraints that rendered items illegal. Therefore, we made sure that the precise consonant-vowel sequence occurred in some real word in the dictionary (i.e., that there was a longer word included the monosyllable as a subsequence). These candidates were presented to one informant, for a final check that they did not mean anything. Together with the 2018 selection, they were recorded using a headset **XXX **mc** please fill in** and an Olympus **XXX **mc** please fill in** from the written form presented together with the same item orally (by the second author). The complete recorded list was finally presented to two more informants, who could repeat all the items and who confirmed there were no real words. Even so, there was one monosyllable that was often identified as a real word (intended "yî" /yXX/; identified as "yi" /yi/, *tree*). This item is removed from analyses below.


The final list is composed of three practice items; 20 monoysllables containing sounds that are less frequent in the world's languages than singleton plosives; 8 bisyllables; 12 trisyllables; and 4 quadrisyllables (see Table 2).

```{r tab2-stims, eval=F}
#add orthographic representations targets & other stims chars
read.delim("stimuli.txt", encoding="UTF-8")->stims
stims=stims[order(stims[,1]),-3]
colnames(stims)<-c("Type","Orthographic","Phonological")
cbindX(stims[stims$Type=="practice",],
      stims[stims$Type==1,],
      stims[stims$Type==2,],
      stims[stims$Type==3,],
      stims[stims$Type==4,])

stims=stims[,-1]

#this is leading to an error
#mytab2 = kable(stims, "latex", booktabs = T,caption="NWR stimuli in orthographic and phonological representations.", row.names =F) %>%
#kable_styling() %>%
#add_header_above(c("Practice" = 2, "Monosyll" = 2, "Bisyll" = 2, "Trisyll" = 2, "Tetrasyll" = 2))

#now that I added phonological representations, even this basic table chokes!
apa_table(stims,"NWR stimuli in orthographic and phonological representations.")
```


A Praat script was written to randomize this list 20 times, and split it into two sublists, to generate 40 different elicitation sets. The 40 elicitation sets are available online from https://osf.io/5qspb. The split had the following constraints:

- the same three items were selected as practice items and used in all 40 elicitation sets
- splits were done within each length group from the 2018 items (i.e., separately for 2, 3, and 4-syllable items); and among onset groups for the difficult monosyllables generated in 2019 (i.e., all the monosyllables starting with tp were split into 2 sublists). Since some of these groups had an odd number of items, one of the sublists was slightly longer than the other (20 versus 23).
- once the sublist split had been done, items were randomized such that all children heard first the 3 practice items in a fixed order (1, 2, and 4 syllables), a randomized version of their sublist selection of difficult onset items, and randomized versions of their 2-syllables, then 3-syllables, and finally 4-syllable items.



### Procedure

We tried to balance three desiderata: That children would not be unduly exposed to the items before they themselves had to repeat them; that children would feel comfortable doing this task with us; and that the community would feel safe with us doing this task with their children. Moreover, there were also some logistic constraints in terms of the space availability. As a result, the places where elicitation happened varied across the hamlets.

We visited four different hamlets once, and attempted to test all eligible children present at the time, to prevent the items 'spreading' through hearsay. In the first hamlet, we tested children in five different places, with some children being tested inside a house and others tested on the veranda. The complete list of places and the ways in which they met the desiderata mentioned above can be found in the raw data, available from online supplementary materials.
 <!--M2A: I didn't look for the SM yet, but probably best not to provide exact hamlet names-->


The child was donned a headset (**xx **mc** please fill in** for most of the children, SHURE WH20 XLR headset with a dynamic microphone for the rest), recorded into the left channel into a Tascam DR40x digital recorder. For most children, the headset could not stay comfortably on the child's head, and thus it was placed on the child's shoulders, with the microphone carefully placed close to the child's mouth. A local informant sat next to the child, to would provide the instructions and, if needed, coach the child to make sure, using the three practice items as well as real words, that they understood that the task was to repeat the items precisely without changing anything. An experimenter (the first author) delivered the elicitation stimuli to the local informant and the child over headphones. 


The first phase was making sure the child understood the task. This was explained orally and the first training item was presented. Often, children froze and did not say anything. If this happened, then we followed this procedure. First, the informant insisted. If the child still did not say anything, the informant asked the child to repeat a real word, and another, and another. If the child could repeat these correctly, then we provided the recorded training item over headphones again. Most children successfully started repeating the items presented over headphones at this point; a few further needed the local informant to model the behavior (i.e., they would hear the item again, and she would say it; then we would play it again, and ask the child to say it). A small minority still failed to repeat the item after hearing it over headphones. If that occurred, we tried with the second training item, at which point some children got it and could continue. A small minority, however, failed to repeat this one, as well as the third training item, in which case we stopped the test altogether. 
<!--M2A: do you know how many these were? oh! it's below. then maybe we can just put a (see Participants below), etc.-->

NWR studies vary in whether children are provided with several opportunities to hear and say the item. To have a fixed and clear procedure, we decided that items other than the inital three training ones would not be repeated unless the child made an attempt to produce them. If this attempt was judged correct by the local informant, then the experimenter would move on to the next item (whispering this over a separate headset that was recording onto the right track of the same Tascam). If the local informant heard a deviation, she indicated to the experimenter that the item needed to be repeated, and up to 5 attempts were allowed. 

Whenever siblings from the same family were tested, an attempt was made to test first the older and then the younger child, and always on different elicitaiton sets.

### Coding

A script was written to randomize all tokens from all children, pairing each with the auditory target the child had been provided with.  A native research assistant then listened through all productions of a given target (randomized across children and repetition order), and made a judgment of whether the item was correctly or incorrectly repeated. She additionally orthographically transcribed exactly what the child said, providing some examples of the types of errors children in general make (without making specific reference to Yélî sounds or the items in the elicitation sets). 

### Analyses

Some NWR studies employ phoneme-based scoring in addition to or instead of word-level accuracy [e.g., @cristia2020infant]. We scored items in terms of the number of phonemes that could be aligned across the target and attempt, divided by the number of phonemes of whichever item was longer (the target or the attempt). Although previous work does not use distance metrics, we additionally report those.
<!--M2A: A concrete scoring example would be a good idea here-->

Additionally, we looked up each phoneme in each target word in the Phoible database [@phoible]. We entered the number of languages in which each phoneme was found, as well as what percentage of languages in the database that represented. We could not find a small number of phonemes in the database, and treat those as NA values that do not contribute to the target-level global average phone cross-linguistic frequency, although an alternative would be to attribute them a very small number.
<!--
A2M: do you agree with this decision or shall we do the alternative?
M2A: yeah, i think adding abirtrarily small frequencies to the NA cases is fraught for the reasons we talked about this morning (i.e., possible systematic undersampling in phoible that can't be helped).
-->

Finally, for describing children's patterns of errors, all repetitions of a given target were taken into account. We describe the proportion of items where the change resulted in a real word (semantic errors); and classify the most common phonological errors.

### Participants

This study was approved as part of a larger research effort by **@ac add info from Middy's FAIR reply**. Participation was voluntary, with children being invited to come and participate. Regardless of how they performed, children were provided with a snack as compensation. Children who came up to participate but then refused were nonetheless provided with the snack. 

```{r demo}
# read.csv("NWR-test-log_FINAL.csv",header=T,sep=";")->fulldemo
# fulldemo=fulldemo[,-2] #remove children's names
# fulldemo$dob=as.Date(as.character(fulldemo$DOB.ISO),format="%Y%m%d")
# fulldemo$test=as.Date(as.character(fulldemo$Date),format="%Y%m%d")
# fulldemo$age.rounded=as.numeric(difftime(fulldemo$test,fulldemo$dob,unit="weeks"))/52.25
# fulldemo$Age=as.numeric(as.character(fulldemo$Age))
# plot(fulldemo$Age~fulldemo$age.rounded)
# write.csv(fulldemo,"nwr-demo.csv")
                            
read.csv("NWR-demo.csv",header=T)->demo
#colnames(demo)

#if family is missing, attribute each child to a different one
demo$familyID=as.character(demo$familyID)
demo$familyID[is.na(demo$familyID)]<-paste("k",demo$ID[is.na(demo$familyID)])

subset(demo,included=="yes" & Age <12 & !is.na(Age))->inc
#if age.rounded is missing, use reported age
inc$age.rounded[is.na(inc$age.rounded)]<-inc$Age[is.na(inc$age.rounded)]
#summary(inc)
table(inc$birthOrder)->mytab
```

A total `r dim(demo)[1]` children were tested, from `r length(levels(factor(demo$familyID)))` of families, in five hamlets. Some children could not be included for the following reasons: refused participation or failed to repeat items presented over headphones even after coaching (N=`r sum(demo$included=="didn\xd5t repeat")`), spoke too softly to allow offline coding (N=`r sum(demo$included=="spoke too softly")`). In addition, `r sum(demo$Age >12 | demo$included=="too old",na.rm=T)` teenagers were tested to put younger children at ease; their data is not included in analyses below. The remaining `r dim(inc)[1]` children (`r sum(inc$Sex=="G")` girls) were aged `r round(mean(inc$age.rounded),2)` years (range `r round(min(inc$age.rounded),2)`-`r round(max(inc$age.rounded),2)` years). There were `r sum(inc$monolingual=="yes", na.rm=T)` children exposed only to Yélî in the home, `r sum(inc$monolingual=="no", na.rm=T)` children who were also exposed to another language in the home, and , `r sum(is.na(inc$monolingual))` for whom this information was missing. Maternal years of education averaged `r round(mean(inc$matEd,na.rm=T),2)` years (range `r round(min(inc$matEd,na.rm=T),2)`-`r round(max(inc$matEd,na.rm=T),2)` years; `r sum(is.na(inc$matEd))` children had this information missing).[^1] In terms of birth order, `r mytab[1]` were first borns, `r mytab[2]` second,  `r mytab[3]` third,  `r mytab[4]` forth,  `r mytab[5]` fifth, `r mytab[6]` sixth, and  `r mytab[7]` did not have this information.

[^1]: Education is often reported in even years because people typically complete two-year cycles.

## Results

```{r wrangling,eval=REDOALL}
source("wrangling.R")

```

### Preliminary analyses


```{r readin}
read.table("final_data.txt",header=T,encoding="UTF-8")->trnsc

#quick fix, but should be fixed upstream @ac
trnsc$Sex[trnsc$Sex=="F"]<-"G"
trnsc$Sex=factor(trnsc$Sex)
trnsc$birthOrder=as.numeric(trnsc$birthOrder)
```


```{r Fig1-first_vs_others, ig.pos = "T!", fig.width=5, fig.height=5, fig.cap="NWR scores for individual participants averaging separately their first attempts and all other attempts."}

aggregate(trnsc$correct*100,by=list(trnsc$id,trnsc$age.rounded,trnsc$attempt),mean)->pwcm_att
colnames(pwcm_att)<-c("ID","Age","Attempt","Word")

#remove children who do not have both first & subsequent attempts
# which is
excl=sum(table(pwcm_att$ID)!=2)

pwcm_att[pwcm_att$ID %in% names(table(pwcm_att$ID))[table(pwcm_att$ID)==2],]->pwcm_att

myx=as.numeric(as.factor(pwcm_att$ID))/100+as.numeric(as.factor(pwcm_att$Attempt))

plot(pwcm_att$Word~myx,pch=20,cex=2,lty=3,xlab="",xaxt="n",ylab="NWR scores (%)")
for(id in levels(factor(pwcm_att$ID))) lines(myx[pwcm_att$ID==id],pwcm_att$Word[pwcm_att$ID==id],lty=2)
axis(1,at=1:2,labels=c("first","others"))

t.test(pwcm_att$Word~pwcm_att$Attempt,paired=T)->myt
```

We first checked whether accuracy varies between first and subsequent presentations of an item by averaging word-level accuracy at the participant level separately for first attempts and subsequent repetitions. We excluded `r excl` children who did not have data for one of these two types. As shown in Figure 1, participants' mean word-level accuracy became more heterogeneous in subsequent repetitions. Surprisingly, subsequent repetitions (M = `r round(mean(pwcm_att$Word[pwcm_att$Attempt!="first"]))`, SD = `r round(sd(pwcm_att$Word[pwcm_att$Attempt!="first"]))`)
were on average less accurate than first ones (M = `r round(mean(pwcm_att$Word[pwcm_att$Attempt=="first"]))`, SD = `r round(sd(pwcm_att$Word[pwcm_att$Attempt=="first"]))`), t(`r round(myt$parameter,2)`) = `r round(myt$statistic,2)`, p = `r round(myt$p.value,2)`. Given the uncertainty in whether previous work used only the first or all repetitions, and since behavior degraded and became more heterogeneous in subsequent repetitions, the rest of the analyses focus on only the first repetitions.

```{r exclude subsequent}
trnsc=trnsc[trnsc$attempt=="first",]

```


```{r overall}

aggregate(trnsc$correct*100,by=list(trnsc$id,trnsc$Sex,trnsc$age.rounded,trnsc$birthOrder,trnsc$matEd),mean)->overall
aggregate(trnsc$pc*100,by=list(trnsc$id,trnsc$Sex,trnsc$age.rounded,trnsc$birthOrder,trnsc$matEd),mean)->pc
aggregate(trnsc$nld*100,by=list(trnsc$id,trnsc$Sex,trnsc$age.rounded,trnsc$birthOrder,trnsc$matEd),mean)->nld
overall=cbind(overall,pc$x,nld$x)
colnames(overall)<-c("ID","Sex","Age","BrthO","matEd","Word","Phoneme","NLD")

```


Taking into account only the first attempts, we averaged attempts by each of the `r dim(overall)[1]` children who had data for first attempts; their ages ranged from `r min(overall$Age)` to `r max(overall$Age)` (M = `r round(mean(overall$Age),1)`, SD = `r round(sd(overall$Age),1)`). 

The overall NWR score was M = `r round(mean(overall$Word))`% (SD = `r round(sd(overall$Word))`%). Scores based on phonemes are even higher  M = `r round(mean(overall$Phoneme))`% (SD = `r round(sd(overall$Phoneme))`%). The phoneme-based normalized Levenshtein distance was  M = `r round(mean(overall$NLD))`% (SD = `r round(sd(overall$NLD))`%), meaning that about a fifth of phonemes were were substituted, inserted, or deleted. Notice that the normalized Levenshtein distance is the complement of phoneme-based scores.



### NWR as a function of cross-linguistic phone frequency

In this analysis, we were interested in variation in performance as a function of the average frequency with which sounds composing individual target words are found in languages over the world. To look at this, we fit a mixed logistic regression, in which the outcome variable was whether the non-word was correctly repeated or not. The fixed effects of interest was the average cross-linguistic phone frequency; we also included child age as a control fixed effect, and child ID & target ID as random variables.

```{r xling-freq-ana}
xl_fr=glmer(correct~avg_fr+age.rounded+(1|file) + (1|target),data=trnsc,family="binomial")
mycoeff=round(summary(xl_fr)$coefficients,3)
```


We could include 885 observations (**why -- shouldn't it be 1k??**), from 41 children producing in any given trial one of 41 potential target words. The analysis revealed a significant estimate for the average cross-linguistic frequency of phones in the target words (ß = `r mycoeff["avg_fr","Estimate"]`, SE ß = `r mycoeff["avg_fr","Std. Error"]`, p = `r mycoeff["avg_fr","Pr(>|z|)"]`): Target words with phones found more frequently across languages had a higher proportion of words that were correctly repeated, as clear in Figure 2. 


```{r Fig2-xling_freq, ig.pos = "T!", fig.width=5, fig.height=5, fig.cap="NWR scores for individual target words as a function of the average frequency with which each phone is found across languages."}
aggregate(trnsc$correct*100,by=list(trnsc$target,trnsc$avg_fr),mean)->xling
colnames(xling)<-c("Target","Avg_fr","Word")
plot(xling$Word~xling$Avg_fr,lty=3,ylab="NWR scores (% words)",xlab="Average phone cross-ling frequency",type="n")
text(xling$Word~xling$Avg_fr,labels=xling$Target,cex=.5)
abline(lm(xling$Word~xling$Avg_fr))
```


```{r diffic,echo=F}
nasalVs=c(":a", ":â" ,":ê", ":i" ,":o")
oralVs=c("a", "â","e",  "é" ,"ê", "i","î" ,"o","ó","u")
simplexCs=c("m","n","l","w","y","v","d","g","p","t","k","f","h","ch")
complexCs=c("dp","tp","kp","km","kn","mb","gh","lv")

allVC=c(nasalVs,oralVs,simplexCs,complexCs)


read.table("substitution_bank.txt",header=T)->sub_bank
colnames(sub_bank)<-c("target","realization")
sub_bank$target=factor(sub_bank$target,levels=allVC)
sub_bank$realization=factor(sub_bank$realization,levels=allVC)

sub_tab=table(sub_bank)[allVC,allVC]


read.table("deletion_bank.txt",header=T)->del_bank
colnames(del_bank)<-c("target")
del_bank$target=factor(del_bank$target,levels=allVC)
del_tab= table(del_bank)[allVC]

#remove English meanings
trnsc$meaning[grep("English",trnsc$meaning)]<-NA
```

Next, we investigated patterns of deletion and substitution. Deletions were relatively rare, with only `r sum(del_tab[c(oralVs,nasalVs)])` vowels deleted, and `r sum(del_tab[c(simplexCs,complexCs)])` consonants. 

As for substitutions, it was as common for a nasal vowel to be produced as an oral vowel as vice versa (`r sum(sub_tab[oralVs,nasalVs])` oral target vowels produced as nasal vowels, `r sum(sub_tab[nasalVs,oralVs])` nasal target vowels produced as oral vowels). Substitutions in which the oral nature was preserved but the quality of the vowel was changed were a great deal more common than changes in quality among nasal vowels (`r sum(sub_tab[oralVs,oralVs])` oral vowels produced with a different quality; `r sum(sub_tab[nasalVs,nasalVs])` nasal vowels produced with a different quality). As for consonants, asymmetries were very marked with more complex consonants (specifically `r paste0(complexCs)`) mispronounced as simple ones (specifically `r paste0(simplexCs)`, `r sum(sub_tab[complexCs,simplexCs])`) than vice versa (`r sum(sub_tab[simplexCs,complexCs])`). Simple consonants were mispronounced as other simple consonants quite frequently (`r sum(sub_tab[simplexCs,simplexCs])` simple consonants mispronounced as other simple ones, compared to `r sum(sub_tab[complexCs,complexCs])` complex ones).

**probably turn this into a table; use proportion of correctly produced rather than number; use first attempts rather than the whole data set**


Finally, we looked into the frequency with which mispronunciations resulted in real words. Nearly all of them were: `r round(sum(!is.na(trnsc$meaning))/sum(!is.na(trnsc$mispronunciation))*100)`%.

** this is higher than any previous report! can we be sure of it?**

### Accuracy a function of item length 

```{r bylength}
aggregate(trnsc$correct*100,by=list(trnsc$id,trnsc$age.rounded,trnsc$type),mean)->by_length
aggregate(trnsc$pc*100,by=list(trnsc$id,trnsc$age.rounded,trnsc$type),mean)->pc
aggregate(trnsc$nld*100,by=list(trnsc$id,trnsc$age.rounded,trnsc$type),mean)->nld
by_length=cbind(by_length,pc$x,nld$x)
colnames(by_length)<-c("ID","Age","Length","Word","Phoneme","NLD")

```


Next, we inspected whether accuracy varied as a function of word length. Results are shown on table XX. This table shows that monosyllables accuracy was much lower than other lengths. This is likely due to the fact that the majority of monosyllables included were chosen because they had sounds that are rare in the world's languages, which may indicate that they are hard to produce or to perceive. 

Setting monosyllables aside, we observe the typical pattern of decreased accuracy for longer items, although this is particularly salient for the whole word scoring. This is the most commonly reported type of score, but it is also the least forgiving. The pattern is less marked when other two scores are used, which are less sensitive to errors.


```{r tablength}
mytab=NULL
for(myvar in c("Word","Phoneme","NLD")){
  mytab=cbind(mytab,
              paste0(round(aggregate(by_length[,myvar],by=list(by_length$Length),mean)$x),
              " (",round(aggregate(by_length[,myvar],by=list(by_length$Length),sd)$x),")"))
}
rownames(mytab)<-levels(by_length$Length)
colnames(mytab)<-c("Word","Phoneme","NLD")
apa_table(mytab,caption="NWR scores measured in whole word accuracy, phoneme-based accuracy, and normalized Levenshtein Distance, separately for the four stimuli lengths.")
```





### Factor structuring individual variation 



```{r Fig2-scores by age, ig.pos = "T!", fig.width=5, fig.height=5, fig.cap="NWR whole-word scores for individual participants as a function of age and sex (blue = boys, pink = girls)."}

plot(overall$Word~overall$Age,pch=20,cex=2,lty=3,ylab="NWR scores (% words)",xlab="Age (years)",col=ifelse(overall$Sex=="B","blue","pink"))

cor.test(overall$Word,overall$Age,method="s")->corage
t.test(overall$Word~overall$Sex)->tsex
cor.test(overall$Word,overall$BrthO,method="s")->corbo
cor.test(overall$Word,overall$matEd,method="s")->corme

```


Our final exploratory analysis assessed whether variation in scores was structured by factors that vary across individuals. As shown in Figure 2, there was a greater deal of variance in earlier than later ages, with significantly higher NWR scores for older children  (Spearman's rank correlation, given inequality of variance, rho (`r corage$statistic`) = `r corage$estimate`, p = `r corage$p.value`). In contrast, there was no clear association between NWR scores, on the one hand, and sex (t (`r tsex$statistic`) = `r tsex$parameter`, p = `r tsex$p.value`), birth order  (data missing for `r sum(is.na(demo$birthOrder[demo$included=="yes"]))` children, `r corbo$statistic`) = `r corbo$estimate`, p = `r corbo$p.value`) or maternal education  (data missing for `r sum(is.na(demo$matEd[demo$included=="yes"]))` children, `r corme$statistic`) = `r corme$estimate`, p = `r corme$p.value`)




## Discussion

- What is the overall repetition accuracy (whole word, phoneme based, distance)?

- How does this change as a function of item complexity (number of syllables, sound complexity)?

1. Children are more accurate for mono-syllables than longer items
1. The length distribution in Yélî words is more balanced than that in English, and thus the performance decline for poly- versus mono-syllables may be less pronounced than that for English. **Check for work on European languages that may have looked into this**


2. Similarly, we do not know of NWR research that manipulates the difficulty of the sounds that are included in the items, but word naming and other research suggests that children are more accurate when producing easy and/or typologically common sounds than difficult and/or typologically rare sounds [CITE]. Therefore, we expect higher accuracy for items with common sounds than in those with rare sounds.
2. The Yélî sound inventory is very large and compressed, with many similar sounds that are acoustic and articulatory neighbors. Therefore, this may constitute a pressure for children to have finer auditory skills (and perhaps more precise articulations) than children speaking languages with a simpler inventory. As a result, differences between easier and harder items may be smaller in this work than in other research. **no work looking at consonants & vowels? no work looking at nasal vowels in particular?**

**(MC: but we can try and do a cursory analysis based on the corpora we have from Steve and my transcription of naturalistic interactions!)**


- How frequent are errors that result in real words? Is that a function of item complexity?
- Is individual variation explainable by child age, sex, birth order, monolingual status, and/or parental education?


3. Children's accuracy increases with child age.
4. Non-monolingual Yélî children are less accurate than monolingual ones when tested on the society-dominant language (we did not test any non-dominant language) 
<!-- **(MC: or just because they're acquiring two phonologies and so show slightly delayed linguistic development, in parallel with vocab development stuff; we don't have data on their dominant language, but i guess we can assume it's YD for most kids---i'm sure there are exceptions!)** ac2mc I added "society-" because that's what I meant (we didn't get proportions of time in which they hear YD versus other languages); I didn't add anything about why this would come out like this or not here, but I'll add this -- although I think the evidence on that is mixed (ie no strong evidence that learning 2 phonologies sets you back)-->
5. As revious NWR evidence on this is mixed, but general findings on language development suggest that children whose mothers are more educated are more accurate than children whose mothers are less educated.
6. To our knowledge, there is no previous NWR work on this, but other research suggests that first-born children should outperform later-born children

4. Anecdotally Yélî children grow up in close-knitted communities and thus may receive significant portions of their language input from people not in their nuclear family 
<!-- **(MC: definitely almost all kin; 5k of them)** ac2mc I'm not sure this is how bio anthropologists think about kin; because it's about how you attribute resources to boost your genes' chances of being passed on relative to those of others in the same space. So my guess is that there is still a difference between nuclear family and not. Rewrote kin -> nuclear family -->
(or at least from people other than their mothers, who tend to be the non-native speakers). If so, the difference between monolinguals and not monolinguals may be smaller than that found in other work
<!-- **(MC: maybe it'd be more useful to try and estimate the number of non-Y moms who live in each hamlet where testing is taking place?)** ac2mc not sure it's worth the effort, in any case other female adults are almost certain to have very little impact on the target child... Perhaps I'm not saying something very clearly? -->
. That said, one recent study on the same population shows that most child-directed input in the first 2 years does come from the mother 
<!-- **(MC: i can get you actual estimates for this for YD; let's go with that instead!)** -->
<!-- ac2mc I think I meant your own paper! added "on the same population "-->
, so in so far as this input has a crucial formational role, then there may still be a performance gap between these two groups.
5. In the Rossel community, formal education plays an extremely minor role in ensuring individual's success, is not a good index of relative socio-economic status, and furthermore there is only a narrow range of variation in maternal educational attainment. This may lead to no or only very small advantages for children whose mothers are more educated, provided that the causal chain between maternal education and child language is via SES more broadly. However, if education directly boosts maternal verbal skills and the incidence of verbal behavior (as suggested by CITE), then we should still see a difference along this factor.
6. One main causal path between birth order and language development is via parental input (CITE). Given our arguments above for how mothers may not be as important among Rossel people than in other places, then the performance gap between first borns and later borns may be smaller.



\newpage

## Acknowledgments

We are grateful to informants and individuals who participated in the study. AC acknowledges financial and institutional support from Agence Nationale de la Recherche (ANR-17-CE28-0007 LangAge, ANR-16-DATA-0004 ACLEW, ANR-14-CE30-0003 MechELex, ANR-17-EURE-0017) and the J. S. McDonnell Foundation Understanding Human Cognition Scholar Award. MC blabla NWO Veni Innovational Scheme Grant (XXX-XX-XXX).


# References

\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}
