---
title             : "Supplementary materials to: Non-word repetition in children learning Yélî Dnye"
shorttitle        : "SM: NWR in Yélî Dnye learners"

author: 
  - name          : "Alejandrina Cristia"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "29, rue d’Ulm, 75005 Paris, France"
    email         : "alecristia@gmail.com"
  - name          : "Marisa Casillas"
    affiliation   : "2,3"
    address       : "Wundtlaan 1, 6525 XD, Nijmegen, The Netherlands"
    email         : "marisa.casillas@mpi.nl"

affiliation:
  - id            : "1"
    institution   : "Laboratoire de Sciences Cognitives et de Psycholinguistique, Département d'Etudes Cognitives, ENS, EHESS, CNRS, PSL University"
  - id            : "2"
    institution   : "Max Planck Institute for Psycholinguistics"
  - id            : "3"
    institution   : "University of Chicago"


author_note: >
       Both authors contributed to study funding, design, data collection, annotation, analyses, writing.

  
abstract: >
   This is a supplementary material containing: 1. Explanation of a small-scale systematic review carried out to embed the main results in the NWR literature; 2. The use of this data to look at potential length effects on NWR as a function of language characteristics; 3. Length and age effects in this and previous work.



bibliography      : ["nwrbib.bib"]

figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : no

lang              : "english"
class             : "doc"
output            :
  papaja::apa6_pdf
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
library(dplyr)
library(openxlsx)
```




## SM1: Small systematic review of previous NWR work 

In order to embed our research questions in the context of previous work, we build our current study on a systematic review of NWR done for a different paper: @cristia2020infant. We preferred to build on a systematic review because, while they are common in psychology research, *unsystematic reviews*—e.g. starting out with familiar papers and then reading other papers citing those; searching via Google Scholar or another search engine; or reading someone else's review—do not generalize and represent biased samples of the literature [@thomas2020impact]. For instance, famous studies are more likely to be found, even though there are no quality differences between more versus less famous studies. 

We here build on a recent and highly relevant systematic review: @cristia2020infant "combined our previous knowledge of the literature and systematic searching to yield a sample of 17 studies that we could interrogate further". Specifically, the main goal of that systematic review was to check whether "effects of infant-directed input quantities may have been reported. Input variability has not been studied, to our knowledge, but two potential proxies of input have: Monolingual status, and socio-economic status.  We set aside the literature comparing monolinguals against non-monolinguals because we reasoned that some readers may argue this contrast does not only show effects of input differences, but also interference across the languages being learned" (Cristia, 2020, personal records). Discovery combined an initial list produced by Gianmatteo Farabolini (based on readings of literature on bilingual NWR mainly), with a scholar.google.com search carried out in incognito mode by Cristia, with keywords "non-word-repetition socio-economic-status". Studies finally included reported on children who were monolingual and between 3 and 7 years of age, and which used proportion of non-words repeated as the outcome metric.

We revisited this selection of articles in order to extract data on actual performance, taking into account also the length of the non-words used in the study. From that selection, only 8 articles reported on actual performance in the text, figures, or tables, fortunately representing a range of languages: Persian by @farmani2018normalization; Israeli Arabic  [@jabere2018xperiment];  English  [@vance2005speech]; Slovak by @kapalkova2013non and @polivsenska2014improving; Sotho by @wilsenach2013phonological; Swedish by @kalnak2014nonword and @radeborg2006swedish. We extended that selection with 5 additional papers selected to increase the number of languages represented, with both languages sometimes described as having short words [Cantonese: @stokes2006nonword; Mandarin:  @lei2011developmental]; and others described as having long words [Italian: @piazzalunga2019articulatory; Spanish: @balladares2016socio]. Finally, we also included the Tsimane' data presented in that paper [@cristia2020infant].


## SM2: Assessing potential length effects 



```{r demo}
# please note that you will not be able to do this because you don't have this version, that is not fully de-identified
# read.csv("NWR-test-log_FINAL.csv",header=T,sep=";")->fulldemo
# fulldemo=fulldemo[,-2] #remove children's names
# fulldemo$dob=as.Date(as.character(fulldemo$DOB.ISO),format="%Y%m%d")
# fulldemo$test=as.Date(as.character(fulldemo$Date),format="%Y%m%d")
# fulldemo$age.rounded=as.numeric(difftime(fulldemo$test,fulldemo$dob,unit="weeks"))/52.25
# fulldemo$Age=as.numeric(as.character(fulldemo$Age))
# plot(fulldemo$Age~fulldemo$age.rounded)
# write.csv(fulldemo,"nwr-demo.csv")

# read.csv("nwr-demo.csv",header=T)->demo
# demo=demo[,-c(20,21)] #remove DOBs 
# write.csv(demo,"nwr-demo.csv")

read.csv("nwr-demo.csv",header=T)->demo
#colnames(demo)


subset(demo,included=="yes" & Age <12 & !is.na(Age))->inc
#if age.rounded is missing, use reported age
inc$age.rounded[is.na(inc$age.rounded)]<-inc$Age[is.na(inc$age.rounded)]
#summary(inc)
table(inc$birthOrder)->birthorder.tab


```


```{r readin}
read.table("final_data.txt",header=T,encoding="UTF-8")->trnsc

trnsc$Sex=factor(trnsc$Sex)
trnsc$birthOrder=as.numeric(trnsc$birthOrder)

# Include only included children in trnsc
trnsc[which(trnsc$id %in% inc$ID),]->trnsc.inc

```


```{r exclude subsequent}
trnsc.inc-> trnsc.inc_all #keep a copy for figure comparing first & subsequent

trnsc.inc=trnsc.inc[trnsc.inc$attempt=="first",]

```

```{r prep-prevlit, echo=F}
pl<- read.csv("previous_lit.csv")

pl_stacked=cbind(pl[,c("Study","Language","n","Age")],stack(pl[c("X1","X2","X3","X4")]),stack(pl[c("sd1","sd2","sd3","sd4")]))
colnames(pl_stacked)[5:8]<-c("Median","Length","SD","sdtype")
pl_stacked$Length=gsub("X","",pl_stacked$Length)

#remove lines with NA median
pl_stacked[!is.na(pl_stacked$Median),]->pl_stacked


#add YD data
aggregate(trnsc.inc$correct*100,by=list(trnsc.inc$id,trnsc.inc$type,trnsc.inc$age.rounded),mean,na.rm=T)->m_ty_id_age #get means by child and length and age

age_cuts=c(3,6,8,13)
m_ty_id_age$age_g=cut(m_ty_id_age$Group.3,age_cuts) #apply cuts to define groups
m_ty_id_age$age_gm<-ifelse(m_ty_id_age$age_g=="(3,6]",4.5,ifelse(m_ty_id_age$age_g=="(6,8]",7,10))


aggregate(m_ty_id_age$x,by=list(m_ty_id_age$Group.2,m_ty_id_age$age_gm),median,na.rm=T)->yd_res #get avg across children within age groups
yd_res=cbind(yd_res,aggregate(m_ty_id_age$x,by=list(m_ty_id_age$Group.2,m_ty_id_age$age_gm),sd,na.rm=T)$x,aggregate(m_ty_id_age$x,by=list(m_ty_id_age$Group.2,m_ty_id_age$age_gm),length)$x) #add sd & ns

colnames(yd_res)<-c("Length","Age","Median","SD","n")
yd_res_toadd=cbind("This paper","Yélî Dnye",yd_res[,c("n","Age","Median","Length","SD")],NA)
colnames(yd_res_toadd)<-c("Study","Language","n","Age","Median","Length","SD","sdtype")

#combine
pl_stacked=rbind(pl_stacked,yd_res_toadd)

#add SE
pl_stacked$se=pl_stacked$SD/sqrt(pl_stacked$n)

#remove monosyllables
pl_stacked=pl_stacked[pl_stacked$Length!=1,]

pl_stacked$Language=factor(pl_stacked$Language)


pl$dif<-pl$X2-pl$X3
difs=aggregate(pl$dif,by=list(pl$Language),mean, na.rm=T)
```



We looked at the subset of papers that reported NWR scores separately for different word lengths. These were: Israeli Arabic  [@jabere2018xperiment]; Cantonese  [@stokes2006nonword]; English  [@vance2005speech]; Italian  [@piazzalunga2019articulatory]; and Tsimane'  [@cristia2020infant]. 


Our reading of that work is that, although there is cross-linguistic (or cross-sample) variation in length effects, these do not systematically line up with expected typical word length in different languages. For instance, the difference in NWR scores for 2- versus 3-syllable items (averaging across age groups) is largest in Tsimane' (~`r round(difs[difs$Group.1=="Tsimane'","x"])`%) and Arabic (~`r round(difs[difs$Group.1=="Arabic","x"])`%), which tend to have longer words, as does Italian, where the difference between 2- and 3-syllable items was only ~`r round(difs[difs$Group.1=="Italian","x"])`%. Similarly, two languages that are often described as heavily biased towards monosyllables show diverse length effects (Cantonese ~`r round(difs[difs$Group.1=="Cantonese","x"])`% versus English ~`r round(difs[difs$Group.1=="English","x"])`%).




```{r fig-prevlit, echo = FALSE, ig.pos = "T!", fig.width=8, fig.height=5.5, fig.cap = "NWR scores as a function of age (in years) and item length for comparable studies (2-4 indicating number of syllables,  2=dashed, 3=dotted, 4=dotted and dashed). Jaber-Awida (2018) reported on 20 Israeli Arabic learners (orange); Piazzalunga et al. (2019) reported on groups of 24-60 Italian learners (black); Stokes et al. (2006) on 15 Cantonese learners (blue); Vance et al. (2005) on 17-20 English learners (light green); Cristia et al. (2020) reported on groups of 4-6 Tsimane' learners (dark green); the present study reports on groups of 8-19 Yélî Dnye learners (purple). Central tendency is the mean except for Italian and Yélî Dnye (median); error is one standard error. Age has been slightly shifted for ease of inspection of different lengths at a given age."}



mycols = c( 
             "Arabic" = "orange",
             "Italian"   ="black",
             "Cantonese"   ="blue",
             "English"   ="green",
            "Tsimane'" = "darkgreen",
             "Yélî Dnye" = "purple")

mywd = c( 
             "Arabic" = 1,
             "Italian"   =1,
             "Cantonese"   =1,
             "English"   =1,
             "Tsimane'" = 2,
             "Yélî Dnye" = 4)

#just a trick to jitter age
jit_age=(pl_stacked$Age+(as.numeric(as.character(pl_stacked$Length))-2)/5)

plot(1,xlab="Average age (years)",ylab="NWR scores (words)",ylim=c(0,100),xlim=range(pl_stacked$Age))


for(thislang in levels(pl_stacked$Language)) for(thislength in levels(factor(pl_stacked$Length[pl_stacked$Language== thislang]))){
  myx=pl_stacked$Age[pl_stacked$Language==thislang & pl_stacked$Length==thislength]
  myx=myx+(as.numeric(as.character(thislength))-2)/5 #jitter the length lines a little bit with respect to avg age
  points(pl_stacked$Median[pl_stacked$Language==thislang & pl_stacked$Length==thislength]~myx,col=mycols[thislang],pch=as.character(thislength))
  myy=pl_stacked$Median[pl_stacked$Language==thislang & pl_stacked$Length==thislength]
  lines(myy~myx, col=mycols[thislang],lty=as.numeric(thislength),lwd=mywd[thislang]) #join points of same length & lang
  mylb=myy-pl_stacked$se[pl_stacked$Language==thislang & pl_stacked$Length==thislength]
  myhb=myy+pl_stacked$se[pl_stacked$Language==thislang & pl_stacked$Length==thislength]
  for(i in 1:length(mylb)) {
    mylty=as.numeric(as.character(thislength))
    lines(rep(myx[i],2),c(mylb[i],myhb[i]),col=mycols[thislang],lty=mylty) #error bar
    lines(c(myx[i]-.1,myx[i]+.1),c(mylb[i],mylb[i]),col=mycols[thislang],lty=mylty) #lower error bar tip
    lines(c(myx[i]-.1,myx[i]+.1),c(myhb[i],myhb[i]),col=mycols[thislang],lty=mylty) #higher error bar tip
    }
} 



```



```{r prep-pl}

pl_overall=pl[,1:7]

#add YD data
yd_res_toadd[yd_res_toadd$Length!=1,]->yd_res_toadd
means=aggregate(yd_res_toadd$Median,by=list(yd_res_toadd$Study,yd_res_toadd$Language,yd_res_toadd$n,yd_res_toadd$Age),mean)
sds=aggregate(yd_res_toadd$SD,by=list(yd_res_toadd$Study,yd_res_toadd$Language,yd_res_toadd$n,yd_res_toadd$Age),mean)
yd_res_toadd2=cbind(means,sds$x)

colnames(yd_res_toadd2)<-c("Study","Language","n","Age","Overall","SD")
yd_res_toadd2$Lengths<-"2-4"
yd_res_toadd2=yd_res_toadd2[,c("Study","Language","n","Age","Lengths","Overall","SD")]

#combine
pl_overall=rbind(pl_overall,yd_res_toadd2)

pl_overall$Language<-gsub(" (.*)","",pl_overall$Language)


read.csv("lang-char.tsv",sep="\t")->lang
merge(pl_overall,lang,by="Language")->pl_overall

#add SE
pl_overall$se=pl_overall$SD/sqrt(pl_overall$n)

#pairs(pl_overall[c("Overall","nchar","phones","words")])

#cor.test(pl_overall$Overall,pl_overall$nchar)
#t.test(pl_overall$Overall~pl_overall$words)

all_lang=levels(factor(pl_overall$Language))

```

## SM3: Integrating our data with that from other studies

For this analysis, we could include all studies that reported non-word repetition scores based on whole item scoring for at least some length, ideally separating children by age. Specifically, Arabic was represented by @jabere2018xperiment; Cantonese by @stokes2006nonword; English by @vance2005speech; Italian by @piazzalunga2019articulatory; Mandarin by @lei2011developmental; Spanish by @balladares2016socio;  Tsimane' by @cristia2020infant; and Yélî Dnye from the present study. Studies varied in the length of non-words that were considered; whenever results were reported separately for different lengths, we calculated overall averages based on lengths of 2 and 3 syllables, for increased comparability. Results separating different age groups are shown in Figure \@ref(fig:fig-prevlit-overall).


```{r fig-prevlit-overall, echo = FALSE, ig.pos = "T!", fig.width=8, fig.height=5.5, fig.cap = "NWR scores as a function of age (in years), averaged across multiple non-word lengths, as a function of children's native languages. The legend indicates language and the length of non-words (in syllables). Central tendency is mean; error is one standard error."}
mycols = c( 
             "Arabic" = "orange",
             "Cantonese"   ="blue",
             "English"   ="green",
             "Italian"   ="black",
             "Mandarin"   ="gray",
             "Persian"   ="pink",
             "Slovak"   ="lightblue",
            "Sotho"   ="magenta",
            "Spanish"   ="brown",
            "Swedish"   ="red",
            "Tsimane'" = "darkgreen",
             "Yélî" = "purple")

mywd = c( 
             "Arabic" = 1,
             "Cantonese"   =1,
             "English"   =1,
             "Italian"   =1,
             "Mandarin"   =1,
             "Persian"   =1,
             "Slovak"   =1,
            "Sotho"   =1,
            "Spanish"   =1,
            "Swedish"   =1,
             "Tsimane'" = 2,
             "Yélî" = 4)

plot(pl_overall$Overall~pl_overall$Age,col=mycols[pl_overall$Language],pch=1,
     xlab="Average age (years)",ylab="NWR scores (words)",ylim=c(0,100))

for(thislang in levels(factor(pl_overall$Language))) {
  myx=pl_overall$Age[pl_overall$Language==thislang ]
  myy=pl_overall$Overall[pl_overall$Language==thislang]
  lines(myy~myx, col=mycols[thislang],lwd=mywd[thislang]) #join points of same length & lang
  mylb=myy-pl_overall$se[pl_overall$Language==thislang]
  myhb=myy+pl_overall$se[pl_overall$Language==thislang]
  for(i in 1:length(mylb)) {
    
    lines(rep(myx[i],2),c(mylb[i],myhb[i]),col=mycols[thislang]) #error bar
    lines(c(myx[i]-.1,myx[i]+.1),c(mylb[i],mylb[i]),col=mycols[thislang]) #lower error bar tip
    lines(c(myx[i]-.1,myx[i]+.1),c(myhb[i],myhb[i]),col=mycols[thislang]) #higher error bar tip
    }
} 

#add legend
#first half of langs
xmin=6
ymax=30
ymin=0
myy=seq(from=ymax,to=ymin,length.out=length(all_lang)/2)
names(myy)<-all_lang[1:(length(all_lang)/2)]
for(thislang in names(myy)){
  lines(c(xmin-.1,xmin+.1),c(myy[thislang],myy[thislang]),col=mycols[thislang]) #line sample
  text(xmin+.15,myy[thislang],paste(thislang,pl_overall$Lengths[pl_overall$Language==thislang][1]),col=mycols[thislang],pos=4)
} 

#second half
xmin=9.5
names(myy)<-all_lang[(length(all_lang)/2+1):length(all_lang)]
for(thislang in names(myy)){
  lines(c(xmin-.1,xmin+.1),c(myy[thislang],myy[thislang]),col=mycols[thislang]) #line sample
  text(xmin+.15,myy[thislang],paste(thislang,pl_overall$Lengths[pl_overall$Language==thislang][1]),col=mycols[thislang],pos=4)
} 
```


Several observations can be drawn from this figure. To begin with, we focus on the comparison between Yélî Dnye and Tsimane'. These two groups have been described as having (very) roughly similar lifestyles and levels of child-directed speech, yet they exhibit very different results: Tsimane' shows lower overall NWR scores (and according to Figure  \@ref(fig:fig-prevlit), larger length effects). What can we conclude from this apparent difference?

One possible lens of interpretation relates NWR differences to the minor differences in children's directed linguistic input rates. However, the differences in methods used to estimate these rates render this first interpretation unreliable. Based on current estimates Tsimane' children may encounter somewhere between similar to somewhat lower rates of directed speech comared to Yélî children. The lower boundary for Tsimane' input rate is established by @cristia2019child, who used behavioral observations of child-inclusive one-on-one conversation. The upper boundary is set by @scaff2021daylong, who used human annotation to detect speech, feeding this into an automated temporal method for assigning speech as child-directed or not (NB: this method could lead to over- but not under-estimation because any nearby speech, e.g. from a female adult, that coincided with child vocalization would count as child-directed). Yélî Dnye estimates come from @casillas2021early, who hand-coded speech with the help of a native research assistant, and then summed all child-directed speech, effectively establishing an upper boundary of all the speech children could potentially process. There is no equivalent 'lower bound' for Yélî Dnye to compare with the Tsimane' findings. So, while it is tempting to link this difference in NWR scores to directed input rates, we will first need to actually establish input rates  using comparable methods for Yéli and Tsimane' children.

Prevalence of literacy is a second possible way to interpret these differences in NWR scores. @cristia2020infant point out the relatively low prevalence of literacy, and more generally, the variable access to formal education available in the Tsimane' population. This is very different from the present Yélî population, where most adults have accumulated some schooling, with many having basic to fluent literacy in English (note: literacy in Yélî Dnye is much less common, especially among the very young; see main text). Under this second hypothesis, there are phonetic effects of learning to read that have significant consequences for young children's encoding and decoding of sounds in the context of NWR tasks. Notice that this is not the same as the oft-recorded effect of learning to read affecting NWR performance illustrated, for instance, in the data for Sotho in Figure \@ref(fig:fig-prevlit-overall). Those two data points have been gathered from two groups of children, all exposed mainly to Sotho, but children with higher NWR had been learning to read in Sotho, whereas those with lower scores were learning to read in English. What is at stake in our proposed alternative interpretation of the lower scores observed among the Tsimane' is related  to literacy *in the broader population* (rather than in the tested children themselves). However, there is at least one more plausible explanation: design and execution of the NWR task

A third, plausible, explanation for these differences in NWR is that the Tsimane' results are not generally comparable to the previous body of literature, and specifically not comparable to our present study. @cristia2020infant administered the NWR in the form of a group game played outside, with a non-native experimenter providing the target, and each person of the group attempting it in their stead. This implies a number of important methodological differences with the standard implementation of NWR, where children are tested individually, they hear items spoken by a native speaker (often over headphones), the experimenter tends to belong to the same community as the children, and testing occurs in quiet conditions (with little background noise). Thus, a priority is for additional data to be gathered using this more novel testing paradigm in other populations; or more from the Tsimane' population, but using the more traditional paradigm.

Broadening our discussion to all of the studies in our literature review, we notice that there is rather wide variation of the range of NWR scores found across these samples, and that, in fact, the strength of age effects also varies. We performed some exploratory analyses to see whether features of the languages children were learning could be related to their overall NWR scores. We extracted the estimated number of phonemes in the language from PHOIBLE and coded whether words in the language tended to be relatively short or long, based on information in the papers cited above and other sources. Neither of these two predictors explained the variance in average NWR scores, illustrated in Figure \@ref(fig:fig-prevlit-overall). It is possible that average word length plays a role, but that researchers tend to incorporate this into their design by including longer items when the native language allows it, with e.g. Sotho non-words having 4--7 syllables in length.

# References

\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}
